---
title: "June2021_analyses"
author: "Elizabeth Forbes"
date: "5/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The files are imported/cleaned/created in the other doc ("TOC_2017incs_ForbesMoore.rmd") in this folder.  Here is data exploration and analyses based on feedback from a May 2021 committee meeting.

The chunk is to analyze the TOC data for possible treatment effects: does treatment have an effect at all, within certain soil types, or is the main driver of effect soil type?
- I would expect treatment to have some effect because increased carbon inputs into the soil from litter that isn't eaten by herbivores (e.g. O > MW, O > C)
- so let's look at ANOVAs of each soil type against treatment for NPOC, TN, and flushes.

one-way ANOVAs to detect treatment effects on soil DOC:
```{r treatmenteffects_NPOCnonfume}
library(tidyverse)
library(rstatix) #package for tidy stats

# one-way ANOVA of open soil samples, DOC:
toc %>% 
  filter(feature == "OS") %>% 
  anova_test(NPOC_nonfume ~ trtmnt) # no difference in one-way anova for OS soils
#   Effect DFn DFd     F     p p<.05   ges
# 1 trtmnt   3 116 0.556 0.645       0.014

# one-way ANOVA of under tree samples, DOC:
toc %>% 
  filter(feature == "UT") %>% 
  anova_test(NPOC_nonfume ~ trtmnt) # no difference in one-way anova for UT soils

# one-way ANOVA of on mound samples, DOC:
toc %>% 
  filter(feature == "OM") %>% 
  anova_test(NPOC_nonfume ~ trtmnt) # no difference in one-way anova for OM soils
```

one-way ANOVAs to detect treatment effects on microbial biomass C:
```{r treatmenteffects_NPOC_flush}
# one-way ANOVA of open soil samples, microbial C:
toc %>% 
  filter(feature == "OS") %>% 
  anova_test(NPOC_flush ~ trtmnt) # no difference in one-way anova for OS soils
#   Effect DFn DFd     F     p p<.05   ges
# 1 trtmnt   3 116 0.556 0.645       0.014

# one-way ANOVA of under tree samples, microbial C:
toc %>% 
  filter(feature == "UT") %>% 
  anova_test(NPOC_flush ~ trtmnt) # no difference in one-way anova for UT soils, p <0.09 tho

# one-way ANOVA of on mound samples, microbial C:
toc %>% 
  filter(feature == "OM") %>% 
  anova_test(NPOC_flush ~ trtmnt) # no difference in one-way anova for OM soils
```
one-way ANOVAs to detect treatment effects on soil DON:
```{r treatmenteffects_TN_nonfume}
# one-way ANOVA of open soil samples, soil DON:
toc %>% 
  filter(feature == "OS") %>% 
  anova_test(TN_nonfume ~ trtmnt) # no difference in one-way anova for OS soils
#   Effect DFn DFd     F     p p<.05   ges
# 1 trtmnt   3 116 0.556 0.645       0.014

# one-way ANOVA of under tree samples, soil DON:
toc %>% 
  filter(feature == "UT") %>% 
  anova_test(TN_nonfume ~ trtmnt) # no difference in one-way anova for UT soils, p <0.09 tho

# one-way ANOVA of on mound samples, soil DON:
toc %>% 
  filter(feature == "OM") %>% 
  anova_test(TN_nonfume ~ trtmnt) # no difference in one-way anova for OM soils
```

one-way ANOVAs to detect treatment effects on soil microbial N:
```{r treatmenteffects_TN_nonfume}
# one-way ANOVA of open soil samples, soil DON:
toc %>% 
  filter(feature == "OS") %>% 
  anova_test(TN_flush ~ trtmnt) # no difference in one-way anova for OS soils
#   Effect DFn DFd     F     p p<.05   ges
# 1 trtmnt   3 116 0.556 0.645       0.014

# one-way ANOVA of under tree samples, soil DON:
toc %>% 
  filter(feature == "UT") %>% 
  anova_test(TN_flush ~ trtmnt) # no difference in one-way anova for UT soils, p <0.09 tho

# one-way ANOVA of on mound samples, soil DON:
toc %>% 
  filter(feature == "OM") %>% 
  anova_test(TN_flush ~ trtmnt) # no difference in one-way anova for OM soils
```

Conclusion: there does not appear to be hints of effect of treatment on soil DOC, DON, microbial biomass C, or microbial biomass N.

Look at distribution of the data, one variable at a time. First, NPOC soil (DOC):
```{r}
library(fitdistrplus)

# remove na's for each of the four cols of interest:
NPOCsoil_naomit <- na.omit(toc$NPOC_nonfume)
NPOCflush_naomit <- na.omit(toc$NPOC_flush)
TNsoil_naomit <- na.omit(toc$TN_nonfume)
TNflush_naomit <- na.omit(toc$TN_flush)

# Cullen and Frey plot, NPOC soil, plus bootstrapping:
descdist(NPOCsoil_naomit, boot=500, discrete = FALSE)
# kurtosis (pointiness) and skew are plotted as blue point ("observation"); possible gamma, beta, weibull distributions

# plot these distributions on the data to see what fits:
fw <- fitdist(NPOCsoil_naomit, "weibull")
fg <- fitdist(NPOCsoil_naomit, "gamma")
par(mfrow = c(2, 2))
plot.legend <- c("Weibull", "gamma")
denscomp(list(fw, fg), legendtext = plot.legend)
qqcomp(list(fw, fg), legendtext = plot.legend)
cdfcomp(list(fw, fg), legendtext = plot.legend)
ppcomp(list(fw, fg), legendtext = plot.legend)

# most likely seeing gamma distributed data for soil DOC

# larger plot, just those two to determine best fit:
denscomp(list(fw, fg), legendtext = c("weibull", "gamma"))
# honestly looks like Weibull is a better fit!  weird.

# and some descriptive statistics, including AIC:
gofstat(list(fw, fg))
# GAMMA is ever-so-slightly a better fit according to AIC, BIC, and goodness-of-fit stats.
```
... and determining data distribution for NPOC flush aka microbial biomass carbon:
```{r}
# Cullen and Frey plot, NPOC flush (plus bootstrapping):
descdist(as.numeric(NPOCflush_naomit), boot=500, discrete = FALSE)
# possible lognormal distribution, gamma distribution, exponential distribution

# plot these distributions on the data to see what fits:
fl <- fitdist(as.numeric(NPOCflush_naomit), "lnorm")
fg <- fitdist(as.numeric(NPOCflush_naomit), "gamma")
fe <- fitdist(as.numeric(NPOCflush_naomit), "exp")
par(mfrow = c(2, 2))
plot.legend <- c("lognormal", "gamma", "exponential")
denscomp(list(fl, fg, fe), legendtext = plot.legend)
qqcomp(list(fl, fg, fe), legendtext = plot.legend)
cdfcomp(list(fl, fg, fe), legendtext = plot.legend)
ppcomp(list(fl, fg, fe), legendtext = plot.legend)

# based on these plots, could be either lognormal or gamma, but I'm leaning gamma.

# larger plot, just those two to determine best fit:
denscomp(list(fl, fg), legendtext = c("lognormal", "gamma"))
# honestly looks like lognormal is a biiiit better of a fit but sort of a toss-up

# and some descriptive statistics, including AIC:
gofstat(list(fl, fg))
# GAMMA is ever-so-slightly a better fit according to AIC, BIC (tho the goodness-of-fit stats somehwat indicate lognormal...(are lower than gamma's)
```

... and determining data distribution for soil TN aka DON:
```{r}
# Cullen and Frey plot, TN soil (plus bootstrapping):
descdist(TNsoil_naomit, boot=500, discrete = FALSE)
# possible gamma distribution, weibull distribution

# plot these distributions on the data to see what fits:
fw <- fitdist(TNsoil_naomit, "weibull")
fg <- fitdist(TNsoil_naomit, "gamma")
fe <- fitdist(TNsoil_naomit, "exp")
par(mfrow = c(2, 2))
plot.legend <- c("weibull", "gamma", "exponential")
denscomp(list(fw, fg, fe), legendtext = plot.legend)
qqcomp(list(fw, fg, fe), legendtext = plot.legend)
cdfcomp(list(fw, fg, fe), legendtext = plot.legend)
ppcomp(list(fw, fg, fe), legendtext = plot.legend)

# based on these plots I'm leaning gamma. Eliminating exponential but wil compare with Weibull.

# larger plot, just those two to determine best fit:
denscomp(list(fw, fg), legendtext = c("weibull", "gamma"))
# looks to me like gamma is a better fit distribution; confirm with desc. stats:

gofstat(list(fw, fg))
# GAMMA is ever-so-slightly a better fit according to AIC, BIC, and goodness-of-fit stats
```

... lastly, determining data distribution for TN flush aka microbial biomass nitrogen:
```{r}
# Cullen and Frey plot, TN flush (plus bootstrapping):
descdist(as.numeric(TNflush_naomit), boot=500, discrete = FALSE)
# possible gamma distribution, weibull distribution

# plot these distributions on the data to see what fits:
fw <- fitdist(as.numeric(TNflush_naomit), "weibull")
fg <- fitdist(as.numeric(TNflush_naomit), "gamma")
fe <- fitdist(as.numeric(TNflush_naomit), "exp")
par(mfrow = c(2, 2))
plot.legend <- c("weibull", "gamma", "exponential")
denscomp(list(fw, fg, fe), legendtext = plot.legend)
qqcomp(list(fw, fg, fe), legendtext = plot.legend)
cdfcomp(list(fw, fg, fe), legendtext = plot.legend)
ppcomp(list(fw, fg, fe), legendtext = plot.legend)

# based on these plots I'm leaning gamma, again.  Will test against Weibull.

# larger plot, just those two to determine best fit:
denscomp(list(fw, fg), legendtext = c("weibull", "gamma"))
# looks to me like gamma is a better fit distribution; confirm with desc. stats:

gofstat(list(fw, fg))
# GAMMA is ever-so-slightly a better fit according to AIC, BIC, and goodness-of-fit stats
```
Okay the above indicates that the response variables are all gamma-distributed.  Before assuming anything tho use an LMM and examine the residuals for normality, and other model validation steps (normality, homogeneity, fixed X, indepdence, and model specs).

```{r histogram}
# remove the 'fume' data from this dataframe for visualization:
toc_fourvar <- toc[,-c(5,7)]
# use forecats to reorder data:
toc_fourvar <- toc_fourvar %>% 
  mutate(feature=fct_relevel(feature, "OS", "UT", "OM")) %>% 
  mutate(trtmnt=fct_relevel(trtmnt, "MWC", "MW", "C", "O"))

library(reshape2)
# melt data to long form for histograms:
toc_long <- melt(data = toc_fourvar,
                 id.vars = c("repl","trtmnt","feature","ID"),
                 variable.name = "soil_variable",
                 value.name = "ugC_gsoil")

# histogram of each X value for toc data, all four categories (plus two more bc )
library(hrbrthemes)
toc_long %>% 
  ggplot(aes(x=ugC_gsoil, fill=soil_variable))+
  geom_histogram(alpha=0.6, position = 'identity')+
  theme_ipsum()+
  labs(fill="")+
  facet_wrap(~soil_variable)+
  theme(legend.position = "none")

# looks like normal distribution of microbial C, and gamma distributions of the other three.
```

```{r linearmodel_testing_soilDOC}
# LMMs: https://www.r-bloggers.com/2017/12/linear-mixed-effect-models-in-r/
# library(nlme)
# library(lme4)
### full LINEAR model to examine residuals, soil DOC:
LM_DOC <- lm(NPOC_nonfume ~ trtmnt*feature, data = toc_fourvar)
summary(LM_DOC)
plot(LM_DOC) # possible outlier based on cook's distance?
cooks <- cooks.distance(LM_DOC)
n <- nrow(toc_fourvar)

# plot outliers: not sure what to do about these yet, will consider
{plot(cooks, pch="*", cex=2, main = "influential observations by cook's distance")
abline(h=4/n, col="red")}

# build ou the GLM as a benchmark for comparison with LMMs (same as LM but comparable):
# library(nlme)
GLM_DOC <- gls(NPOC_nonfume ~ trtmnt*feature, data = toc_fourvar, method = "ML")
summary(GLM_DOC)

# LMMs with fixed effects/random effects:
library(lme4)
lmm1 <- lmer(NPOC_nonfume ~ trtmnt*feature + (1 | repl), toc_fourvar)
summary(lmm1) # non-significant interaction between any treatment:feature; drop from model
lmm2 <- lmer(NPOC_nonfume ~ trtmnt+feature + (1|repl), data = toc_fourvar)
summary(lmm2) # O (treatment) is different from MWC; UT and OM different from OS (feature)
lmm3 <- lmer(NPOC_nonfume ~ feature + (1|repl), data = toc_fourvar) # drop treatment entirely just to investigate most parsimonious option
summary(lmm3)

# ANOVA for likelihood-ratio-test between models
anova(lmm2, lmm3) # LMM3 has lower AIC, BIC, and logLik values (p-val is significant in model comparison); but, given we know there might be a treatment effect and that the treatment is literally built into how the KLEE works, go with lmm2 (e.g. including treatment as fixed effect but not with an interaction)

# look at residuals of lmm2:
# QQ plots:
qqnorm(resid(lmm2, type = "pearson")) #woof!  we aren't remotely close to normal.  The outliers? 
hist(resid(lmm2)) # yeah very right-skewed residuals here
plot(lmm2) # also possibly *some* heteroskedasticity (residuals against fitted vals)

# looks like we'll need to do GLMM instead bc the residuals for our best-fitting LMM are not normal.  This is POSSIBLY driven by a handful of outliers.

# but also: try log-transforming your continuous response variable:
lmm4 <- lmer(log(NPOC_nonfume) ~ feature+trtmnt + (1|repl), data = toc_fourvar)
summary(lmm4)
plot(lmm4)
hist(resid(lmm4)) # residuals much more normally distributed, one outlier still.

```
Going to build the same model as above, but in GLMM (generalized linear mixed effects model) which can handle non-normal distributions of data (in our case, as explored above, gamma-distributed) to see if we can tease out trustworthy results without transforming the data (next step if this doesn't work)
```{r soilC_glmm}
# library(lme4)
glmm1 <- glmer(NPOC_nonfume ~ trtmnt*feature + (1|repl),
               data = toc_fourvar,
               family = Gamma,
               control = glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=2e5)))
# model failed to converge, large eigenvalue; ned to check singularity/convergence tolerances on this one
summary(glmm1)

# drop interaction between treatment and feature:
glmm2 <- glmer(NPOC_nonfume ~ trtmnt+feature + (1|repl),
               data = toc_fourvar,
               family = Gamma,
               control = glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=2e5)))
summary(glmm2)

glmm3 <- glmer(NPOC_nonfume ~ feature + (1|repl),
               data = toc_fourvar,
               family = Gamma)
summary(glmm3)

# compare models despite convergence issues:
anova(glmm1, glmm2) # no significant difference between the models, plus glmm2 has lower AIC/BIC values; drop the interaction.
anova(glmm2, glmm3) # hmm!  glmm2 has lower AIC but higher BIC (barely).  significant difference between the two.  but in the interest of bringing my brain to the party, realize that there is a significant difference between MWC and O soils (at least for soil DOC) according to glmm2 so do not drop this fixed term.  Makes it easier to 'choose' glmm2, which includes treatment as a fixed effect with NO interaction.

# adjust link function (default is inverse, and can be an issue according to Ben Bolker) to log link function. This is *possibly* okay considering that I'm already considering log-transforming my data due to the outliers/right-skewness. This is funcitonally the same as a linear model with log-transformed data, so we're comparing arithmetic means here (not geometric mean).
glmm4 <- glmer(NPOC_nonfume ~ trtmnt+feature + (1|repl),
               data = toc_fourvar,
               family = Gamma(link = "log"))
# model converged!
summary(glmm4) # same output; significant differences between O and MWC for treatment, and between OM and UT compared to OS.
anova(glmm2, glmm4) # glmm4 is EVER SO SLIGHTLY a better fit than glmm3.

glmm5 <- glmer(NPOC_nonfume ~ trtmnt+feature + (1|repl),
               data = toc_fourvar,
               family = Gamma(link = "identity"))
# huh!  also converged.
summary(glmm5)
anova(glmm4, glmm5) # glmm 4 still ever so slightly better according to AIC, BIC, loglik

# plot residuals of glmm4:
# The following code demonstrate producing a residual plot on the link scale.
ggplot(data.frame(eta=predict(glmm4,type="link"),
                  pearson=residuals(glmm4,type="pearson")),
      aes(x=eta,y=pearson)) +
    geom_point() +
    theme_bw()
# independent variable by variable
ggplot(data.frame(x1=toc_long$feature,pearson=residuals(glmm4,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw() #hmm some outlier-y behavior in residuals for UT
ggplot(data.frame(x1=toc_long$trtmnt,pearson=residuals(glmm4,type="pearson")),
      aes(x=x1,y=pearson)) +
    geom_point() +
    theme_bw() #outlier residuals in O, MWC
qqnorm(residuals(glmm4))

# histogram of the residuals 
hist(resid(glmm4)) #MORE normally distributed but still less so than log-transformed data (lmm4)
```

So at this point, it's a matter of deciding if I want to go with a linear mixed effects model on log-transformed data or a GLMM on non-transformed data, but with a log link function (e.g. log of means, rather than log of response variable then mean'ed).  So we want to do some digging into our model diagnosics using the package DHARMa, which allows for residual testing for mixed models.
```{r DHARMa}
library(DHARMa)
plotResiduals(glmm4, form = toc_fourvar$feature)
simulateResiduals(glmm4, plot=TRUE)
# non-normally distributed residuals, and have within-group deviations from uniformity e.g. homogeneity of variance is false (heteroskedasticity)
testDispersion(glmm4)
testZeroInflation(glmm4)

plotResiduals(lmm4, form = toc_fourvar$feature)
simulateResiduals(lmm4, plot=TRUE)
testDispersion(lmm4)
# ok so based on these model diagnostics, it looks like log-transformed data is actually the way to go.  some outliers but the residuals look much better.

anova(glmm4, lmm4)
# okay wow yeah.  much lower AIC, BIC, and smaller loglik values.  p-value for the difference between the two is highly significant.

## the following package allows you to get reports/p-values for the LMER objects which do not produce p-values
library(devtools)
# devtools::install_github("easystats/report")
library(report)
report(lmm4)
# report(glmm4)
```

I want to visualize the model outputs: https://lmudge13.github.io/sample_code/mixed_effects.html
```{r visualizemodels}
library(tidyverse) #for all data wrangling
library(cowplot) #for manuscript ready figures
library(lme4) #for lmer & glmer models
# devtools::install_github("strengejacke/sjPlot", force=TRUE) #for plotting lmer and glmer mods
library(sjPlot)
library(sjmisc)
library(effects)
library(sjstats) #use for r2 functions

# make table of model results:
sjPlot::tab_model(lmm4)
sjPlot::tab_model(lmm4,
                  show.re.var = TRUE,
                  pred.labels = c("(Intercept)", "under tree", "termite mound",
                                  "wildlife only", "cattle only", "no herbivores"),
                  dv.labels = "Effects of treatment or soil type on soil DOC (log scale)")
# plot model results too (effect sizes):
sjPlot::plot_model(lmm4,
                   order.terms = c(3, 4, 5, 1, 2), # for some reason this one was in a different order than the others, so manually order here and label in next line
                   axis.labels = c("termite mound", "under tree", "no herbivores", "cattle only", "wildlife only"),
                   show.values = TRUE, show.p = TRUE,
                   title = "Effect of treatment and soil type on soil DOC (log scale)",
                   vline.color = "pink")+
  theme_cowplot()
```


Okay!  time to run models and diagnostics on the other three variables.
```{r soil_DON}
### full LINEAR model to examine residuals, soil DOC:
lm_DON <- lm(TN_nonfume ~ trtmnt*feature, data = toc_fourvar)
summary(lm_DON) # no interaction between treatment and soil type, drop from full model
plot(lm_DON) # outliers, heteroskedasticity, and non-normality of residuals; either glmm or log-transformation

# linear mixed model, full model
lmm1_don <- lmer(TN_nonfume ~ trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm1_don)
sjPlot::tab_model(lmm1_don) # no interaction between treatment and feature, drop

lmm2_don <- lmer(TN_nonfume ~ trtmnt+feature + (1|repl), data=toc_fourvar)
summary(lmm2_don)
sjPlot::tab_model(lmm2_don) # significant effect of O treatment, again.  keep 'trtmnt'.
qqnorm(resid(lmm2_don, type = "pearson")) # still has outliers!

lmm3_don <- lmer(log(TN_nonfume)~trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm3_don)
sjPlot::tab_model(lmm3_don) # drop interaction for log-transformed model

lmm4_don <- lmer(log(TN_nonfume) ~ trtmnt + feature + (1|repl), data=toc_fourvar)
summary(lmm4_don)
sjPlot::tab_model(lmm4_don) 

# DHARMa diagnostics:
plotResiduals(lmm4_don, form = toc_fourvar$feature) #: within-group deviation from uniformity n.s.; Levene Test for homogeneity of variance significant; because we know this, at least across features, we know that variation of residuals across groups are not constant.  Should consider modelling variances, e.g. via ~dispformula in glmmTMB?  but also no significant problems detected across groups in residual vs. predicted plot.  CONFUSED.
simulateResiduals(lmm4_don, plot=TRUE) # some outliers but no problematic points
testDispersion(lmm4_don)

anova(lmm3_don, lmm4_don) #lmm4 (with just main effect not interaction) is a better fit according to AIC, BIC parameters.

# table of model outcomes:
sjPlot::tab_model(lmm4_don)
sjPlot::tab_model(lmm4_don,
                  show.re.var = TRUE,
                  pred.labels = c("(Intercept)", "wildlife only", "cattle only", "no herbivores",
                  "under tree", "termite mound"))

# plot model outcomes:
sjPlot::plot_model(lmm4_don,
                   axis.labels = c("termite mound", "under tree", "no herbivores",
                                   "cattle only", "wildlife only"),
                   show.values = TRUE, show.p = TRUE,
                   title = "Effect of treatment and soil type on soil DON (log scale)",
                   vline.color = "pink")+
  theme_cowplot()
```

Microbial biomass carbon analysis:
```{r micro_DOC}
# full model, no mixed effects
lm_mic <- lm(NPOC_flush ~ trtmnt*feature, data = toc_fourvar)
summary(lm_mic) # no interaction between treatment and soil type, drop from full model
plot(lm_mic) # outliers, heteroskedasticity, and non-normality of residuals; either glmm or log-transformation of data

# full model with mixed effects, interaction between treatment and soil feature:
lmm1_mic <- lmer(NPOC_flush ~ trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm1_mic)
sjPlot::tab_model(lmm1_mic)# no interaction between treatment and feature, drop
qqnorm(resid(lmm1_mic)) # huh just two outliers...
simulateResiduals(lmm1_mic, plot=TRUE) # residuals deviation from normality significant (qq plot) and heteroskedasticity across groups

# full model with mixed effects, interaction between treatment and soil feature, on LOG TRANSFORMED data:
lmm2_mic <- lmer(log(NPOC_flush) ~ trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm2_mic)
sjPlot::tab_model(lmm2_mic)# no interaction between treatment and feature, drop
simulateResiduals(lmm2_mic, plot=TRUE) # residuals normally distributed, no heteroscedasticiy detected across groups

# model with mixed effects, no interaction
lmm3_mic <- lmer(log(NPOC_flush) ~ trtmnt + feature + (1|repl), data=toc_fourvar)
summary(lmm3_mic)
sjPlot::tab_model(lmm3_mic) # no impact of treatment, funnily.  drop to see how iti mpacts things.

simulateResiduals(lmm3_mic, plot=TRUE) # looks alright

# table of model outcomes:
sjPlot::tab_model(lmm3_mic)
sjPlot::tab_model(lmm3_mic,
                  show.re.var = TRUE,
                  pred.labels = c("(Intercept)", "wildlife only", "cattle only", "no herbivores",
                  "under tree", "termite mound"))

# plot model outcomes:
sjPlot::plot_model(lmm3_mic,
                   axis.labels = c("termite mound", "under tree", "no herbivores",
                                   "cattle only", "wildlife only"),
                   show.values = TRUE, show.p = TRUE,
                   title = "Effect of treatment and soil type on soil microbial biomass carbon (log scale)",
                   vline.color = "pink")+
  theme_cowplot()
```
Microbial biomass nitrogen:
```{r micro_DON}
# full model, no mixed effects
lm_mic_tn <- lm(TN_flush ~ trtmnt*feature, data = toc_fourvar)
summary(lm_mic_tn) # no interaction between treatment and soil type, drop from full model
plot(lm_mic_tn) # outliers, heteroskedasticity, and non-normality of residuals; either glmm or log-transformation of data

# full model with mixed effects, interaction between treatment and soil feature:
lmm1_mic_tn <- lmer(TN_flush ~ trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm1_mic_tn)
sjPlot::tab_model(lmm1_mic_tn) # no interaction between treatment and feature, drop
simulateResiduals(lmm1_mic_tn, plot=TRUE) # residuals deviation from normality significant (qq plot) and heteroskedasticity across groups

# full model with mixed effects, interaction between treatment and soil feature, on LOG TRANSFORMED data:
lmm2_mic_tn <- lmer(log(TN_flush) ~ trtmnt*feature + (1|repl), data=toc_fourvar)
summary(lmm2_mic_tn)
sjPlot::tab_model(lmm2_mic_tn) # okay weird; no significant effect at all, of any parameter
simulateResiduals(lmm2_mic_tn, plot=TRUE) # residuals normally distributed, no heteroscedasticiy detected across groups

# model with mixed effects, no interaction
lmm3_mic_tn <- lmer(log(TN_flush) ~ trtmnt + feature + (1|repl), data=toc_fourvar)
summary(lmm3_mic_tn)
sjPlot::tab_model(lmm3_mic_tn) # okay so no impact of treatment, but significant impacts of feature.  feels weird to do this after no effect in previous model?  but had already detected no interaction in non-log-transformed data, so I think this is fine.
simulateResiduals(lmm3_mic_tn, plot = TRUE) # cool; normally distributed residuals, evenly dispersed variance across groups

# table of model outcomes:
sjPlot::tab_model(lmm3_mic_tn)
sjPlot::tab_model(lmm3_mic_tn,
                  show.re.var = TRUE,
                  pred.labels = c("(Intercept)", "wildlife only", "cattle only", "no herbivores",
                  "under tree", "termite mound"))

# plot model outcomes:
sjPlot::plot_model(lmm3_mic_tn,
                   axis.labels = c("termite mound", "under tree", "no herbivores",
                                   "cattle only", "wildlife only"),
                   show.values = TRUE, show.p = TRUE,
                   title = "Effect of treatment and soil type on soil microbial biomass nitrogen (log scale)",
                   vline.color = "pink")+
  theme_cowplot()
```

Next up we want to use the eemeans package to run pairwise comparisons of these models that accounts for multiple comparisons. We're going to stick with comparisons *just* across feature for all four response variables despite the fact that there is some effect of treatment for soil DOC and DON; this is the comparison we're interested in, and we already know that there is no significant difference between treatments other than MWC and O (did try those pairwise comparisons already). (See meeting notes with Josh Sat. July 31st 2021.)
```{r pairwise}
# reminder of models we're keeping:
# lmm3_mic (microbial biomass carbon), lmm3_mic_tn (microbial biomass nitrogen), lmm4_don (soil DON), and lmm4 (soil DOC)
# https://aosmith.rbind.io/2019/03/25/getting-started-with-emmeans/#back-transforming-results

# this command does all pairwise comparisons with the model, and automatically adjusts for multiple comparisons with a Bonferroni correction. Type = "response" allows for the estimates and CI's to be back-transformed from the original log scale, though the test itself is on the model (aka log-transformed) scale. Simple = "each" constrains the pairwise comparisons to those between the fixed effects (treatment and feature).

# below I'm making a test dataframe/linear model for the log-transformed data, because for some reason the function to plot the data (both response and modeled response) is not working with log within the lmer model itself.  Not sure what's up, but since it's just to get the individual points on the plot, we can use an LM of the log-transformed data as below. See https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/issues.html
toc_fourvar$log_NPOC <- log(toc_fourvar$NPOC_nonfume)
testmod <- lmer(log_NPOC ~ feature+trtmnt+(1|repl), data = toc_fourvar)

# calculate estimated marginal means and pairwise comparisons for soil DOC (the real model used here, so we can get the real p-values and confidence intervals in the contrast table below):
emm_soilDOC <- emmeans(lmm4, specs = "feature")

# also make a testmod table for estimated means, for putting points on the plot, see https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/issues.html, since this command doesn't seem to be able to handle models with log-transformed data
emm_testmod <- emmeans(testmod,
                       specs = "feature")

# compute contrasts table of estimated effects with confidence intervals and p-values, using real emmeans table from two steps above
lmm4_simple <- contrast(emm_soilDOC, # use the real model here to calculate pairwise comparisons
                        method = "pairwise",
                        simple = "each",
                        combine = TRUE,
                        type = "response") %>% 
  summary(infer = TRUE)

# plot response: this response has both the modeled means (big points with error bars) and the actual data (log-transformed small jittered)
ggplot_the_response(
  fit = testmod,                 # plots jittered (log-transformed) points on plot
  fit_emm = emm_soilDOC,         # plots model fit points (large points) on plot
  fit_pairs = lmm4_simple,       # plots pairwise comparisons on plot
  palette = pal_okabe_ito_blue,
  y_label = "soil DOC concentration, mg/kg soil (natural log scale)"
)+
  scale_x_discrete(labels=c("OS" = "Open Soil", "UT" = "Under Tree",
                              "OM" = "Termite Mound"))
```
soil DON:
```{r soilDON}
toc_fourvar$log_TN <- log(toc_fourvar$TN_nonfume)
testmod <- lmer(log_TN ~ feature+trtmnt+(1|repl), data = toc_fourvar)

# calculate estimated marginal means and pairwise comparisons for soil DOC (the real model used here, so we can get the real p-values and confidence intervals in the contrast table below):
emm_soilDON <- emmeans(lmm4_don, specs = "feature")

# also make a testmod table for estimated means, for putting points on the plot, see https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/issues.html, since this command doesn't seem to be able to handle models with log-transformed data
emm_testmod <- emmeans(testmod,
                       specs = "feature")

# compute contrasts table of estimated effects with confidence intervals and p-values, using real emmeans table from two steps above
lmm4_simple <- contrast(emm_soilDON, # use the real model here to calculate pairwise comparisons
                        method = "pairwise",
                        simple = "each",
                        combine = TRUE,
                        type = "response") %>% 
  summary(infer = TRUE)

# plot response: this response has both the modeled means (big points with error bars) and the actual data (log-transformed small jittered)
ggplot_the_response(
  fit = testmod,                 # plots jittered (log-transformed) points on plot
  fit_emm = emm_soilDON,         # plots model fit points (large points) on plot
  fit_pairs = lmm4_simple,       # plots pairwise comparisons on plot
  palette = pal_okabe_ito_blue,
  y_label = "soil DON concentration, mg/kg soil (natural log scale)"
)+
  scale_x_discrete(labels=c("OS" = "Open Soil", "UT" = "Under Tree",
                              "OM" = "Termite Mound"))
```
microbial biomass C:
```{r micC}
toc_fourvar$log_micC <- log(toc_fourvar$NPOC_flush)
testmod <- lmer(log_micC ~ feature+trtmnt+(1|repl), data = toc_fourvar)

# calculate estimated marginal means and pairwise comparisons for soil DOC (the real model used here, so we can get the real p-values and confidence intervals in the contrast table below):
emm_micC <- emmeans(lmm3_mic, specs = "feature")

# also make a testmod table for estimated means, for putting points on the plot, see https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/issues.html, since this command doesn't seem to be able to handle models with log-transformed data
emm_testmod <- emmeans(testmod,
                       specs = "feature")

# compute contrasts table of estimated effects with confidence intervals and p-values, using real emmeans table from two steps above
lmm4_simple <- contrast(emm_micC, # use the real model here to calculate pairwise comparisons
                        method = "pairwise",
                        simple = "each",
                        combine = TRUE,
                        type = "response") %>% 
  summary(infer = TRUE)

# plot response: this response has both the modeled means (big points with error bars) and the actual data (log-transformed small jittered)
ggplot_the_response(
  fit = testmod,                 # plots jittered (log-transformed) points on plot
  fit_emm = emm_micC,         # plots model fit points (large points) on plot
  fit_pairs = lmm4_simple,       # plots pairwise comparisons on plot
  palette = pal_okabe_ito_blue,
  y_label = "microbial biomass C, mg/kg soil (natural log scale)"
)+
  scale_x_discrete(labels=c("OS" = "Open Soil", "UT" = "Under Tree",
                              "OM" = "Termite Mound"))
```
microbial biomass N:
```{r micN}
toc_fourvar$log_micN <- log(toc_fourvar$TN_flush)
testmod <- lmer(log_micN ~ feature+trtmnt+(1|repl), data = toc_fourvar)

# calculate estimated marginal means and pairwise comparisons for soil DOC (the real model used here, so we can get the real p-values and confidence intervals in the contrast table below):
emm_micN <- emmeans(lmm3_mic_tn, specs = "feature")

# also make a testmod table for estimated means, for putting points on the plot, see https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/issues.html, since this command doesn't seem to be able to handle models with log-transformed data
emm_testmod <- emmeans(testmod,
                       specs = "feature")

# compute contrasts table of estimated effects with confidence intervals and p-values, using real emmeans table from two steps above
lmm4_simple <- contrast(emm_micN, # use the real model here to calculate pairwise comparisons
                        method = "pairwise",
                        # simple = "each",
                        combine = TRUE,
                        type = "response") %>% 
  summary(infer = TRUE)

# plot response: this response has both the modeled means (big points with error bars) and the actual data (log-transformed small jittered)
ggplot_the_response(
  fit = testmod,                 # plots jittered (log-transformed) points on plot
  fit_emm = emm_micN,         # plots model fit points (large points) on plot
  fit_pairs = lmm4_simple,       # plots pairwise comparisons on plot
  palette = pal_okabe_ito_blue,
  y_label = "microbial biomass N, mg/kg soil (natural log scale)"
)+
  scale_x_discrete(labels=c("OS" = "Open Soil", "UT" = "Under Tree",
                              "OM" = "Termite Mound"))

View(lmm4_simple)
```
## April 15, 2022
# Additional analyses
Exploration of CN ratios
Exploration of total soil carbon (aka both extractable organic carbon and microbial biomass carbon)

FIRST UP: CN ratios of soil microbial communities
```{r}
# use forecats to reorder data:
toc <- toc %>% 
  mutate(feature=fct_relevel(feature, "OS", "UT", "OM")) %>% 
  mutate(trtmnt=fct_relevel(trtmnt, "MWC", "MW", "C", "O"))

library(reshape2)
# melt data to long form for histograms:
toc_long2 <- melt(data = toc,
                 id.vars = c("repl","trtmnt","feature","ID"),
                 variable.name = "soil_variable",
                 value.name = "ugC_gsoil")

# in addition, calculate CN ratio of microbial biomass (higher = fungal dominated, lower = bacteria dominated)
toc$micCN_ratio <- toc$NPOC_flush / toc$TN_flush

# look for outliers!
hist(toc$micCN_ratio, breaks=100) # yes one weirdo outlier way out in the boonies at 100, one at 75, and a handful at 40. Generally speaking I think it'd be ok to eliminate those because soil microbial CN ratios just...can't be that high (and indeed are closer to like, 8:1).
# take out those higher than 25 (there's a break in the data there).
toc_micCN <- toc %>% 
  filter(micCN_ratio < 25)
hist(toc_micCN$micCN_ratio, breaks = 200) # still right-skewed

mypalette <- c("#E69F00", "#009E73", "#D55E00")
toc_micCN %>% 
  ggplot(aes(x=feature, y=log(micCN_ratio), fill = feature))+
  labs(x ="" , y = "log transformed microbial C:N")+
  geom_boxplot()+
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+
  scale_fill_manual("",  
                      labels = c("OS" = "grassland matrix", 
                                 "UT" = "tree canopy",
                                 "OM" = "termite mound"), 
                    values = mypalette)+
                      theme(
  strip.background = element_blank(),
  strip.text.x = element_blank())

```

Let's get down to model testing.
```{r}
# test model:
test_lm <- lm(micCN_ratio ~ trtmnt*feature, data = toc_micCN)
summary(test_lm)
plot(test_lm)
simulateResiduals(test_lm, plot=TRUE)
# data is very non-normally distributed!  outliers and deviation from normal. Within-group deviations of variance

hist(toc$micCN_ratio, breaks=100) # yes one weirdo outlier way out in the boonies at 100, one at 75, and a handful at 40. Generally speaking I think it'd be ok to eliminate those because soil microbial CN ratios just...can't be that high (and indeed are closer to like, 8:1).
# take out those higher than 25 (there's a break in the data there).
toc_micCN <- toc %>% 
  filter(micCN_ratio < 25)
hist(toc_micCN$micCN_ratio, breaks = 200) # still right-skewed

# add random effect aka lme function (nlme package)
test_lme <- lme(micCN_ratio ~ trtmnt*feature, random=~1|repl, data = toc_micCN)
summary(test_lme)
qqnorm(test_lme) #qqplot shows hump in the middle
hist(resid(test_lme)) #residuals are close to normal but a bit right-skewed

# test different distributions aside from normal:
fw <- fitdist(toc_micCN$micCN_ratio, "weibull")
fg <- fitdist(toc_micCN$micCN_ratio, "gamma")
fl <- fitdist(toc_micCN$micCN_ratio, "lnorm")
par(mfrow = c(2, 2))
plot.legend <- c("Weibull", "gamma", "lognormal")
denscomp(list(fw, fg, fl), legendtext = plot.legend) #lognormal distribution looks better
qqcomp(list(fw, fg, fl), legendtext = plot.legend) #again, lognormal
cdfcomp(list(fw, fg, fl), legendtext = plot.legend) # gamma
ppcomp(list(fw, fg, fl), legendtext = plot.legend) #ok, lognormal again.
# so try modeling the distribution with a lognormal distribution, not a normal distribution. In this case you can go with a linear mixed effects model, since we'll be modeling on log-transformed data that thus has a normal distribution. (aka we can leave the generalized linear mixed-effects model for another time.)

# LOG TRANSFORMATION LMM:
test_lmm <- lmer(log(micCN_ratio) ~ feature*trtmnt + (1|repl), 
                 data = toc_micCN)
summary(test_lmm)
plot(test_lmm)
hist(resid(test_lmm)) # residuals much more normally distributed, but not totally
report(test_lmm) # no interaction btw treatment and feature

# drop interaction between treatment, feature
test_lmm <- lmer(log(micCN_ratio) ~ feature+trtmnt + (1|repl), data = toc_micCN)
summary(test_lmm)
simulateResiduals(test_lmm, plot=TRUE)
# library(report) # get p-values for your lmer model 
report(test_lmm) # OM significantly lower CN ratio than soils from UT, OS; no effect of treatment on microbial CN. beta = -0.19, 95% CI [-0.29, -0.09], t(322) = -3.81, p < .001

# also did some model testing with glmer, which does not require normal (gaussian) distribution of the residuals, but the best-fitting model in all those cases was still the lmm with log-transformed data here.

# emmeans to do pairweise comparisons by feature:
emm_cn <- emmeans(test_lmm, specs = "feature")
emm_cn
# contrast table:
p <- contrast(emm_cn, method = "pairwise",
                        simple = "each",
                        combine = TRUE,
                        type = "response") %>% 
  summary(infer = TRUE)
p
```

