---
title: "june2021_analyses"
author: "Elizabeth Forbes"
date: "5/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This doc is for data exploration/analyses related to jar incubations. Data are imported/cleaned in the other RMD doc in this project ("allyrs_concat_jarincs.Rmd"). Inc_b is all the data; inc16-18 are just those years' of data.

The chunk is to analyze the incubation data for possible treatment effects: does treatment have an effect at all, within certain soil types, or is the main driver of effect soil type?
- I would expect treatment to have some effect because increased carbon inputs into the soil from litter that isn't eaten by herbivores (e.g. O > MW, O > C)
- so look at ANOVAs of each soil type against treatment for total accumulation (hr == 144) and repeated measures (aka rate of increase).
```{r treatmenteffects_totacc}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(rstatix) #package for tidy stats

# one-way ANOVA of open soil samples, total C accumulation:
inc_b %>% 
  filter(hrs_incubated == 144) %>%
  filter(soiltype == "OS") %>% 
  anova_test(ugC_gsoil ~ treatment) # significant effect of treatment on OS soils, p = 0.008
# one-way ANOVA of open soil samples, total C accumulation, just 2018:
# removed code but same as above, just added 'year_col' filter; significant effect of treatment in 2018, not 2015 or 2016. See notes from convo with Ana and Devyn about that. 

# one-way ANOVA of under tree samples, total C accumulation:
inc_b %>% 
  filter(hrs_incubated == 144) %>%
  filter(soiltype == "UT") %>% 
  anova_test(ugC_gsoil ~ treatment) # NO significant treatment effect for UT soils, p = 0.34

# one-way ANOVA of termite mound samples, total C accumulation:
inc_b %>% 
  filter(hrs_incubated == 144) %>%
  filter(soiltype == "OM") %>% 
  anova_test(ugC_gsoil ~ treatment) # significant effect of treatment on OM soils, p = 0.014

# conclusion: there is a significant impact of treatment on soil type results, but perhaps only in 2018. 
```

So how does landscape feature influence [carbon metric], and does grazing treatment influence the relative size of this effect? (https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html)
Proposed format:
glmm(C ~ treatment * feature + year + 1|random)
- random effects: treatment nested within block 
- adding year as fixed effect but assuming that year (proxy for drought status) does not interact with treatment/feature on the fundamental level (e.g. carbon) over such a short period


```{r setup}
# model of total accumulation, with interaction between treatment and soil type plus fixed effect of year, and a random effect of treatment nested within location (block):
# using both fixed and random effects, e.g. using GLMMs (generalized linear mixed models)
inc_acc <- inc_b %>% 
  filter(hrs_incubated == 144)
summary(inc_acc)
# make year_col a factor:
inc_acc$year_col <- as.factor(inc_acc$year_col)

# add a rate column: divide "ugC_gsoil" by six (total days) (per notes from Josh, July 31 2021)
inc_acc$ugC_gsoil_day <- (inc_acc$ugC_gsoil)/6

# remove NAs:
inc_acc <- inc_acc[complete.cases(inc_acc[ ,12]),] 

# histogram of rate data:
hist(inc_acc$ugC_gsoil_day) # no major outliers, vaguely gamma-distributed

```

Model testing:
```{r analyses}
lm <- lm(ugC_gsoil_day ~ treatment*soiltype*year_col, data=inc_acc)
anova(lm) # no interaction between treatment and soiltype, but definitely an effect of both fixed

lm1 <- lm(ugC_gsoil_day ~ treatment*year_col + soiltype*year_col , data=inc_acc)
anova(lm1)
hist(resid(lm1)) # not horribly abnormal, but also do want to incorporate the block as a random effect.
# library(DHARMa)
simulateResiduals(lm1, plot = TRUE) # some abnormality in distribution of residuals

# library(lme4)
# library(lmerTest)
# library(emmeans)
lmm_rate <- lmerTest::lmer(ugC_gsoil_day ~ treatment*year_col + soiltype*year_col + (1|replicate), data = inc_acc)
anova(lmm_rate)
summary(lmm_rate)

hist(resid(lmm_rate)) # honestly the residuals don't look too bad.

simulateResiduals(lmm_rate, plot = TRUE) # well apparently they do, booooooo

####################
# log-transformation:
lmm1_rate <- lmerTest::lmer(log(ugC_gsoil_day) ~ treatment*soiltype + 
                              treatment*year_col + 
                              soiltype*year_col +
                              (1|replicate), data=inc_acc)
# library(multcomp)
anova(lmm1_rate) # all three fixed effects (treatment, soil type, and year) are significant; interaction between treatment and year is signficant, as is interaction between soiltype and year

summary(lmm1_rate, test = adjusted("holm")) # use holm-bonferroni correction for multiple testing
hist(resid(lmm1_rate)) # cautiously optimistic
qqnorm(resid(lmm1_rate)) # hey this looks okay!
simulateResiduals(lmm1_rate, plot = TRUE) # huh; better, but not quite there.
plotResiduals(lmm1_rate, form = inc_acc$soiltype) # looks like there is some non-homogeneity of variance within the OM group of data, but across treatments it's okay.  Gonna just be alright with that; when I analyze by year it might be interesting.
anova(lmm1_rate)

inc_acc %>% 
  group_by(year_col) %>% 
  get_summary_stats(ugC_gsoil_day, type = "mean_se")

# what was the effect of drought for pairwise comparisons?
emmeans(lmm1_rate, specs = pairwise ~ "year_col", type = "response")

```
Next up: analyze rate by year. Collected in 2015/incubated in 2016:
```{r analyze_rate_2015}
library(lmerTest)
inc_acc15 <- inc_acc %>% 
  filter(year_col == "2015")
hist(inc_acc15$ugC_gsoil_day, breaks = 100)

lm <- lm(ugC_gsoil_day ~ treatment*soiltype, data=inc_acc15)
anova(lm) # no interaction between treatment and soiltype, but definitely an effect of both fixed

lm1 <- lm(ugC_gsoil_day ~ treatment + soiltype , data=inc_acc15)
anova(lm1)
hist(resid(lm1)) # not horribly abnormal, but also do want to incorporate the block as a random effect.

# library(lme4)
lmm15_rate <- lmerTest::lmer(ugC_gsoil_day ~ treatment + soiltype + (1|replicate), data = inc_acc15)
anova(lmm15_rate)
summary(lmm15_rate)
hist(resid(lmm15_rate))
qqnorm(resid(lmm15_rate))
simulateResiduals(lmm15_rate, plot = TRUE)

# log-transformed data:
lmm15_2_rate <- lmerTest::lmer(log(ugC_gsoil_day)~treatment + soiltype + (1|replicate), data=inc_acc15)
# look at ANOVA table for Type I (sequential type I anova table) from lmerTest package for analyzing LMER objects
anova(lmm15_2_rate) # effect of treatment and not soiltype in 2015!  fascinatinggggg.

summary(lmm15_2_rate, test = adjusted("holm"))
hist(resid(lmm15_2_rate))
simulateResiduals(lmm15_2_rate, plot = TRUE) # this is the model to stick with for 2015 rate data

# tukey's post-hoc tests to determine effects of treatment (tukey is default):
e2015 <- emmeans(lmm15_2_rate,
                 specs = pairwise ~ "treatment", # compare across treatment
                 type = "response")  # convert back from log transformation
e2015$emmeans     # examine modeled means
e2015$contrasts   # examine p-values of pairwise comparisons
```

Collected in 2016/incubated in 2017:
```{r analyze_rate_2016}
inc_acc16 <- inc_acc %>% 
  filter(year_col == "2016")
hist(inc_acc16$ugC_gsoil_day, breaks = 100)

lm <- lm(ugC_gsoil_day ~ treatment*soiltype, data=inc_acc16)
anova(lm) # no interaction, just impact of soiltype (keep treatment in model)

lm1 <- lm(ugC_gsoil_day ~ treatment + soiltype , data=inc_acc16)
anova(lm1)
hist(resid(lm1)) # also do want to incorporate the block as a random effect.

lmm16_rate <- lmerTest::lmer(ugC_gsoil_day ~ treatment + soiltype + (1|replicate), data = inc_acc16)
anova(lmm16_rate)
hist(resid(lmm16_rate))
simulateResiduals(lmm16_rate, plot = TRUE) # almost there, but not quite

lmm16_2_rate <- lmerTest::lmer(log(ugC_gsoil_day) ~ treatment + soiltype + (1|replicate), data=inc_acc16)
# look at ANOVA table for Type I (sequential type I anova table) from lmerTest package for analyzing LMER objects
anova(lmm16_2_rate) # effect of both treatment and soiltype in this model

summary(lmm16_2_rate)
hist(resid(lmm16_2_rate))
simulateResiduals(lmm16_2_rate, plot = TRUE) # okay things look good.  moving on.  both treatment and soiltype effect on soils from 2016.

# tukey's post-hoc tests to determine effects of treatment (tukey is default):
e2016 <- emmeans(lmm16_2_rate,
                 specs = pairwise ~ soiltype|treatment, # compare across treatment + soiltype
                 type = "response")  # convert back from log transformation
e2016$emmeans     # examine modeled means
e2016$contrasts   # examine p-values of pairwise comparisons

# note: did the below comparison with just treatment too, and no individual comparison was statistically significant to p < 0.05 (though MWC/C was close)

# okay within-gropu comparison was illuminating re: feature. Now just do feature:
e2016_2 <- emmeans(lmm16_2_rate,
                   specs = pairwise ~ "soiltype",
                   type = "response")
e2016_2$emmeans
e2016_2$contrasts
```

Collected in 2018/incubated in 2018:
```{r analyze_rate_2016}
inc_acc18 <- inc_acc %>% 
  filter(year_col == "2018")
hist(inc_acc18$ugC_gsoil_day, breaks = 100)

lm <- lm(ugC_gsoil_day ~ treatment*soiltype, data=inc_acc18)
anova(lm) # interaction, and individual impacts of both fixed effects

lmm18_rate <- lmerTest::lmer(ugC_gsoil_day ~ treatment*soiltype + (1|replicate), data = inc_acc18)
anova(lmm18_rate)
hist(resid(lmm18_rate), breaks = 100)
simulateResiduals(lmm18_rate, plot = TRUE) # almost there, but not quite

lmm18_2_rate <- lmerTest::lmer(log(ugC_gsoil_day) ~ treatment*soiltype + (1|replicate), data=inc_acc18)
# look at ANOVA table for Type I (sequential type I anova table) from lmerTest package for analyzing LMER objects
anova(lmm18_2_rate, type=c("I")) # effect of treatment, soiltype, and the interaction in 2018

summary(lmm18_2_rate)
hist(resid(lmm18_2_rate), breaks = 100)
qqnorm(resid(lmm18_2_rate))
simulateResiduals(lmm18_2_rate, plot = TRUE) # Hmm; despite looking okay in terms of qqplot and relatively normal-looking residuals, both DHARMa tests are showing deviations from the norm. Gonna go with this model for now anyway (Aug 3rd, 2021) because it otherwise looks ok.
plotResiduals(lmm18_2_rate)

##########################
# tukey's post-hoc tests to determine effects of treatment (tukey is default):
e2018 <- emmeans(lmm18_2_rate,
                 specs = pairwise ~ soiltype|treatment, # compare across treatment + soiltype
                 type = "response")  # convert back from log transformation
e2018$emmeans     # examine modeled means
e2018$contrasts   # examine p-values of pairwise comparisons

e2018_3 <- emmeans(lmm18_2_rate,
                   specs = pairwise ~ treatment,
                   type = "response")
e2018_3$emmeans
e2018_3$contrasts

# okay within-gropu comparison was illuminating re: feature. Now just do feature:
e2018_2 <- emmeans(lmm18_2_rate,
                   specs = pairwise ~ "soiltype",
                   type = "response")
e2018_2$emmeans
e2018_2$contrasts

```

Interaction plot for 2018:
```{r 2018_interaction}
interaction.plot(
  x.factor = inc_acc18$treatment,
  trace.factor = inc_acc18$soiltype,
  response = inc_acc18$ugC_gsoil_day,
  fun = median,
  ylab = expression(paste("carbon accumulated " (mu*"g/g soil"))),
  xlab = "treatment",
  trace.label = "soil type",
  lwd = 3,
  col = c("#E69F00", "#009E73", "#D55E00")
)

mypalette <- c("#E69F00", "#009E73", "#D55E00")

plot_model(lmm18_2_rate, type = "pred", terms = c("treatment", "soiltype"))+
  theme_classic()+
  scale_color_manual(values = mypalette) #color-blind friendly palette
  

# both methods of producing a plot demonstrate an interaction between treatment and soiltype, but just for OM and OS soils; not UT.
# - OS soils always have lower rates of mineralization than OM *except* in MW plots, where the relationship is switched.
# - no matter what, UT soils have the highest rate of interaction
```

